Let $\{a_n\}$ be a sequence, and define a new sequence $\{s_n\}$ of ``partial sums'' by
$$s_n=\sum_{i=1}^n a_i = a_1 + \cdots + a_n.$$
We write 
$$\lim _{n \to \infty} s_{n}=\lim _{n \to \infty} \sum_{i=1}^{n} a_{i}=\sum_{i=1}^{\infty} a_{i}=\sum a_n$$
and call $\sum a_n$ an \textit{infinite series}. We are interested in when this quantity is finite.
\begin{itemize}
    \item If $\displaystyle\lim _{n \to \infty} s_{n}$ is finite, we say that $\sum a_n$ is \textit{convergent} (or that $\sum a_n$ converges).
    \item  When $\displaystyle\lim _{n \to \infty} s_{n}$ is infinite, we say that $\sum a_n$ is \textit{divergent} (or that $\sum a_n$ diverges).
\end{itemize}

\vspace{.5em}

\begin{multicols}{2}

\begin{tcolorbox}
\textbf{Divergence Test.}
\begin{itemize}
    \item If $\displaystyle \lim_{n\to\infty}a_n\neq 0$, then $\sum a_n$ will diverge.
    \item If $\sum a_n$ converges, then $\displaystyle \lim_{n\to\infty}a_n=0$.
\end{itemize}
\end{tcolorbox}


\begin{tcolorbox}
\textbf{Alternating Series Test.}\newline
Suppose that we have a series $\sum a_{n}$ and either $$a_{n}=(-1)^{n} b_{n}\quad\text{ or }\quad  a_{n}=(-1)^{n+1} b_{n}$$ where $b_{n} \geq 0$ for all $n .$ Then if,
\begin{enumerate}
    \item $\displaystyle\lim _{n \to \infty} b_{n}=0$ and
    \item $\left\{b_{n}\right\}$ is a decreasing sequence
\end{enumerate}
the series $\sum a_{n}$ is convergent.
\end{tcolorbox}

\begin{tcolorbox}
\textbf{Limit Comparison Test.}\newline
Suppose that we have two series $\sum a_n$ and $\sum b_n$ with $a_n\geq 0$ and $b_n > 0$ for all $n$. Let 
$$c=\lim_{n\to\infty}\frac{a_n}{b_n}.$$
If $c$ is positive and finite, then either both series converge or both series diverge.
\end{tcolorbox}


\begin{tcolorbox}
\textbf{Comparison Test.}\newline
Suppose $0\leq a_n\leq b_n$ for all $n$. Then
$$\sum b_n\text{ converges}\implies \sum a_n\text{ converges}$$
and (by contraposition)
$$\sum a_n\text{ diverges}\implies \sum b_n\text{ diverges}.$$
\end{tcolorbox}


\end{multicols}

\begin{tcolorbox}
\textbf{Integral Test.}\newline
Suppose that $f(x)$ is a continuous, positive, and decreasing function on the interval $[k,\infty)$ and that $f(n)=a_n$. Then
$$\int_k^\infty f(x)\ dx\text{ is convergent} \iff \sum_{n=k}^\infty a_n\text{ is convergent}.$$
\end{tcolorbox}
