\documentclass{article}

% set font encoding for PDFLaTeX, XeLaTeX, or LuaTeX
\usepackage{ifxetex,ifluatex}
\if\ifxetex T\else\ifluatex T\else F\fi\fi T%
  \usepackage{fontspec}
\else
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{lmodern}
\fi

\usepackage{hyperref, amssymb, amsmath, multicol}

\usepackage[margin=1in]{geometry}

\pagenumbering{gobble}


\newcommand{\R}{\mathbb{R}}


\renewcommand{\thesubsection}{\Alph{subsection}.\!\!\!\!}


\begin{document}


\begin{center}
\begin{Huge}Integration Application Exercises\end{Huge}
\end{center}


\subsection{Kinematics}
\begin{enumerate}
\item If a particle's position at time $t$ is $x(t)=\cos(t)\sin(t)$, what is the total distance it travels from $t=0$ to $t=2\pi$?
\end{enumerate}


\subsection{Average Values}
\begin{enumerate}
\item Find the average value of the function $f(x) = \cos^3(x)$ on the interval $[0,\pi]$.

\end{enumerate}

\subsection{Probability}

\subsection{Volumes and Surface Area}
\begin{enumerate}
\item Let $R$ be the region above the graph of $x^2+2x-3$ and below the $x$-axis. Find the volume of the solid obtained by rotating the region $R$ around the $x$-axis.
\item A cylindrical hole of radius 3 cm is bored through the center of a wooden sphere of radius 5 cm. Find the volume of the resulting wooden object.
\item Find the volume of the finite region between $x^2-2$ and $2x-2$ revolved around the $x$-axis. Then compute the same for the $y$-axis.
\item Find the surface area and volume of the solid obtained by rotating the graph of $f(x)=\frac{1}{x}$ around the $x$-axis for $x\geq1$.
\item The finite region bounded by the curves $x = 0$, $y = x^2$, and $y = 1$ is rotated about the the $y$-axis. Find the volume of the resulting solid by using the cylindrical shells method.
\item What is the volume of the solid obtained by rotating the region enclosed by $x^2-x$ and $2x^2$ around the line $y=2$.
\item Derive the volume of a hemisphere.
\end{enumerate}



\subsection{Arc Length}

\begin{enumerate}
\item Find the length of the parameterized curve in the plane given by $x=t^2$ and $y=\frac{2}{3}t^3$ with $t\in[0,1]$.
\item Find the length of the curve given by the equation $9y^2=4(x+1)^3$ and the constraints $0 \leq x \leq 1$, $y \geq 0$.
\end{enumerate}


\subsection{Moments and Centroids}

\subsection{Work}

\begin{enumerate}
 \item A 10m-long metal chain of total mass 20 kilograms is lying flat on the floor. How much work would be done if someone grabbed one end of the chain and lifted it 5 meters above the floor? Assume that the acceleration due to gravity is  9.8m/sec$^2$, and that the lifted portion of the chain always hangs vertically.




\end{document}
